---
title: "The exploration's report"
date: "September, 2019"
---

```{r setup_exploration, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos =  "h")
knitr::opts_knit$set(root.dir = "../")

# libraries for spatial data manipulation
library(rmarkdown)
library(dplyr)
library(knitr)
library(rgdal)
library(raster)
library(spdep)
library(bamlss)
library(shapefiles)
```

# Exploration 1

Describe your exploration 1 here, including some plots, tables or any resource used.

```{r exploration_target_1, echo=TRUE, out.width = '100%'}
# your code with the exploration 1 goes here
```

# Exploration 2

Describe your exploration 2 here, including some plots, tables or any resource used.

```{r exploration_target_2, echo=TRUE, out.width = '100%'}
# your code with the exploration 2 goes here
```


# Exploration 3

Describe your exploration 3 here, including some plots, tables or any resource used.

```{r exploration_target_3, echo=TRUE, out.width = '100%'}
# your code with the exploration 3 goes here
```

# Exploration n

Describe your exploration n here, including some plots, tables or any resource used.

```{r exploration_target_n, echo=TRUE, out.width = '100%'}
# your code with the exploration n goes here
```
